
%\clearpagebefore
%\phantomsection
%\addcontentsline{toc}{chapter}{Bibliography}
%\begin{onehalfspacing}
%\bibliographystyle{ieeetr}
%\bibliography{ref}
%\bibliography{reference}
%\end{onehalfspacing}

%\begin{thebibliography}{00}
%\end{thebibliography}

\begin{thebibliography}{00}
\addcontentsline{toc}{chapter}{\bibname}

% 영문저널의 경우
%    \bibitem{ref1} B. Jeon and J. Jeong, ``Blocking artifacts
%    reduction in image compression with block boundary discontiunity
%    criterion,'' {\em IEEE Transactions on Circuits and Systems for
%    Video Tech.}, vol. 8, no.3, pp. 345-357, June 1998.

% 영문학술대회의 경우
%    \bibitem{ref2} W. G. Jeon and Y. S. Cho, ``An equalization
%    technique for OFDM and MC-CDMA in a multipath fading channels,''
%    in {\em Proceedings of IEEE Conference on Acoustics, Speech and
%    Signal Processing}, Munich, Germany, May 1997. pp. 2529-2532.

% 단행본의 경우
%    \bibitem{ref4} C. Mead and L. Conway, {\em Introduction to VLSI
%    Systems}, Addison-Wesley, Boston, 1994.

% URL
%    \bibitem{ref5} The SolarMESH Network,
%    http://owl.mcmater.ca/solarmesh

% Technical Report의 경우
%    \bibitem{ref6} K. E. Elliott and C. M. Greene, ``A local adaptive
%    protocol,'' Argonne National Laboratory, Argonne, France,
%    Technical Report 916-1010-BB, 1997.

% 학위논문의 경우
%    \bibitem{ref7} T. Kim, ``Scheduling and Allocation Problems in
%    High-level Synthesis,'' Ph. D. Dissertation, ECE Department,
%    Univ. of Illinois at U-C, 1993.

% 특허의 경우
%    \bibitem{ref8} Sunghyun Choi, ``Wireless MAC protocol based on a
%    hybrid combination of slot allocation, token passing, and
%    polling for isochronous traffic,'' U.S. Patent No. 6,795,418,
%    September 21, 2004.

% 표준
%    \bibitem{ref9} IEEE Std. 802.11-1999, Part 11: Wireless LAN
%    Medium Access Control (MAC) and Physical Layer (PHY)
%    specifications, Reference number ISO/IEC 8802-11:1999(E), IEEE
%    Std. 802.11, 1999 edition, 1999.


\bibitem{Janus_albrecht}
C.~Albrecht, A.~Merchant, M.~Stokely, M.~Waliji, F.~Labelle, N.~Coehlo, X.~Shi, and C.~E.~Schrock,
``Janus: Optimal Flash Provisioning for Cloud Storage Workloads,"
in \emph{Proceedings of the USENIX Annual Technical Conference} (ATC), 2013.

\bibitem{15nmMLC_Sako}
Mario Sako
Y.~Watanabe, T.~Nakajima, J.~Sato, K.~Muraoka, M.~Fujiu, F.~Kouno, M.~Nakagawa, M.~Masuda, K.~Kato, Y.~Terada, Y.~Shimizu, M.~Honma, A.~Imamoto,  T.~Araya, H.~Konno, T.~Okanaga, T.~Fujimura, X.~Wang, M.~Muramoto, M.~Kamoshida, M.~Kohno, Y.~Suzuki, T.~Hashiguchi, T.~Kobayashi, M.~Yamaoka, and R.~Yamashita,
``A Low-Power 64Gb MLC NAND-Flash Memory in 15 nm CMOS Technology,"
in \emph{Proceedings of the IEEE International Solid-State Circuits Conference} (ISSCC), 2015.

\bibitem{21nmMLC_Kim}
C.~Kim, J.~Ryu, T.~Lee, H.~Kim, J.~Lim, J.~Jeong, S.~Seo, H.~Jeon, B.~Kim, I.~Lee, D.~Lee, P.~Kwak, S.~Cho, Y.~Yim, C.~Cho, W.~Jeong, K.~Park, J.-M.~Han, D.~Song, K.~Kyung, Y.-H.~Lim, and Y.-H.~Jun,
``A 21 nm High Performance 64 Gb MLC NAND Flash Memory with 400 MB/s Asynchronous Toggle DDR Interface,"
\emph{IEEE Journal of Solid-State Circuits}, vol. 47, no. 4, pp. 981--989, 2012.

\bibitem{TLC_Shin}
S.-H.~Shin, D.-K.~Shim, J.-Y.~Jeong, O.-S.~Kwon, S.-Y.~Yoon, M.-H.~Choi, T.-Y.~Kim, H.-W.~Park, H.-J.~Yoon, Y.-S.~Song, Y.-H.~Choi, S.-W.~Shim, Y.-L.~Ahn, K.-T.~Park, J.-M.~Han, K.-H.~Kyung, and Y.-H.~Jun,
``A New 3-bit Programming Algorithm using SLC-to-TLC Migration for 8 MB/s High Performance TLC NAND Flash Memory,"
in \emph{the IEEE Symposium on VLSI Circuits Digest of Technical Papers} (VLSI), 2012.

\bibitem{Flash_Brewer}
J.~E.~Brewer and M.~Gill,
``Nonvolatile Memory Technologies with Emphasis on Flash,"
\emph{Wiley}, 2008.

\bibitem{MooresLaw_chien}
A.~A.~Chien and V.~Karamcheti,
``Moore's Law: The First Ending and a New Beginning,"
\emph{IEEE Computer Magazine}, vol. 46, no. 12, pp. 48--53, 2013.

\bibitem{HotCold_Hsieh}
J.-W. Hsieh, T.-W.~Kuo, and L.-P.~Chang,
``Efficient Identification of Hot Data for Flash Memory Storage Systems,"
\textit{ACM Transactions on Storage}, vol. 2, no. 1, pp. 22--40, 2006.

\bibitem{CAFTL_Chen}
F.~Chen, T.~Luo, and X.~Zhang,
``CAFTL: A Content-Aware Flash Translation Layer Enhancing the Lifespan of Flash Memory Based Solid State Drives,"
in \textit{Proceedings of the USENIX Conference on File and Storage Technologies} (FAST), 2011.

\bibitem{Compression_Lee}
S.~Lee, J.~Park, K.~Fleming, Arvind, and J.~Kim,
``Improving Performance and Lifetime of Solid-State Drives Using Hardware-Accelerated Compression,"
\textit{IEEE Transactions on Consumer Electronics}, vol. 57, no. 4, pp. 1732--1739, 2011.

\bibitem{DT_Lee}
S.~Lee, T.~Kim, K.~Kim, and J.~Kim,
``Lifetime Management of Flash-Based SSDs Using Recovery-Aware Dynamic Throttling,"
in \textit{Proceedings of the USENIX Conference on File and Storage Technologies} (FAST), 2012.

\bibitem{SelfHealing_Wu}
Q.~Wu, G.~Dong, and T.~Zhang,
``Exploiting Heat-Accelerated Flash Memory Wear-Out Recovery to Enable Self-Healing SSDs,"
in \emph{Proceedings of the USENIX Workshop on Hot Topics in Storage and File Systems} (HotStorage), 2011.

\bibitem{RetentionRelaxation_Liu}
R.-S.~Liu, C.-L.~Yang, and W.~Wu,
``Optimizing NAND Flash-Based SSDs via Retention Relaxation,"
in \textit{Proceedings of the USENIX Conference on File and Storage Technologies} (FAST), 2012.

\bibitem{Flexfs_Lee}
S.~Lee, K.~Ha, K.~Zhang, J.~Kim, and J.~Kim,
``FlexFS: A Flexible Flash File System for MLC NAND Flash Memory,"
in \emph{Proceedings of the USENIX Annual Technical Conference} (ATC), 2009.

\bibitem{Cox_JEDEC}
A.~Cox,
``JEDEC SSD Specifications Explained,"
Available: http://www.jedec.org.

\bibitem{DPES_jeong}
J.~Jeong, S.~S.~Hahn, S.~Lee, and J.~Kim,
``Lifetime Improvement of NAND Flash-based Storage Systems Using Dynamic Program and Erase Scaling," 
in \emph{Proceedings of the USENIX Conference on File and Storage Technologies} (FAST), 2014.

\bibitem{BER_mielke}
N.~Mielke, T.~Marquart, N.~Wu, J.~Kessenich, H.~Belgal, E.~Schares, F.~Trivedi, E.~Goodness, and L.~R.~Nevill,
``Bit Error Rate in NAND Flash Memories,"
in \emph{Proceedings of the International Reliability Physics Symposium} (IRPS), 2008.

\bibitem{ReadDisturb_Kang}
M.~Kang, K.-T.~Park, Y.~Song, S.~Hwang, B.~Y.~Choi, Y.~Song, Y.~Lee, and C.~Kim,
``Improving Read Disturb Characteristics by Self-Boosting Read Scheme for Multilevel NAND Flash Memories,"
\emph{Japanese Journal of Applied Physics}, vol. 48, no. 4, pp. 04C062-1--04C062-6, 2009.

\bibitem{ISPP_suh}
K.-D.~Suh, B.-H.~Suh, Y.-H.~Lim, J.-K.~Kim, Y.-J.~Choi, Y.-N.~Koh, S.-S.~Lee, S.-C.~Kwon, B.-S.~Choi, J.-S.~Yum, J.-H.~Choi, J.-R.~Kim, and H.-K.~Lim,
``A 3.3 V 32 Mb NAND Flash Memory with Incremental Step Pulse Programing Scheme,"
\emph{IEEE Journal of Solid-State Circuits}, vol. 30, no. 11, pp. 1149--1156, 1995.

\bibitem{FN_schuegraf}
K.~F.~Schuegraf and C.~Hu,
``Effects of Temperature and Defects on Breakdown Lifetime of Thin SiO2 at Very Low Voltages,"
\emph{IEEE Transactions on Electron Devices}, vol. 41, no. 7, pp. 1227--1232, 1994.

\bibitem{TR_jeong}
J.~Jeong and J.~Kim,
``Dynamic Program and Erase Scaling in NAND Flash-based Storage Systems,"
\emph{Technical Report, Seoul National University},
cares.snu.ac.kr/download/TR-CARES-01-14, 2014.

\bibitem{Reliability_cho}
S.~Cho,
``Improving NAND Flash Memory Reliability with SSD Controllers,"
in \emph{Proceedings of the Flash Memory Summit}, 2013.

\bibitem{JEDEC}
JEDEC standard,
``Stress-Test-Driven Qualification of Integrated Circuits,"
JESD47H.01, 2011.

\bibitem{ISPE_Lee}
D.~W.~Lee, S.~Cho, B.~W.~Kang, S.~Park, B.~Park, M.~K.~Cho, K.-O.~Ahn, Y.~S.~Yang, and S.~W.~Park,
``The Operation Algorithm for Improving the Reliability of TLC (Triple Level Cell) NAND Flash Characteristics,"
in \emph{Proceedings of the IEEE International Memory Workshop} (IMW), 2011.

\bibitem{Flashbench_lee}
S.~Lee, J.~Park, and J.~Kim,
``FlashBench: A Workbench for a Rapid Development of Flash-based Storage Devices,"
in \emph{Proceedings of the IEEE International Symposium on Rapid System Prototyping} (RSP), 2012.

\bibitem{Revisiting_Jung}
M.~Jung and M.~Kandermir,
``Revisiting Widely Held SSD Expectations and Rethinking System-Level Implications,"
in \emph{Proceedings of the ACM SIGMETRICS}, 2013.

\bibitem{WL_Chang}
L.-P. Chang,
``On Efficient Wear Leveling for Large-Scale Flash-Memory Storage Systems,"
in \textit{Proceedings of the ACM Symposium on Applied Computing}, 2007.

\bibitem{ReadRetry_Yang}
J. Yang,
``High-Efficiency SSD for Reliable Data Storage Systems,"
in \textit{Proceedings of the Flash Memory Summit}, 2011.

\bibitem{Reliability_Frickey}
R. Frickey,
``Data Integrity on 20 nm NAND SSDs,"
in \textit{Proceedings of the Flash Memory Summit}, 2012.

\bibitem{Hrtimer}
IBM,
``Kernel APIs, Part 3: Timers and Lists in the 2.6 Kernel,"
http://www.ibm.com/developerworks/library/l-timers-list/.

\bibitem{Narayanan2008}
D. Narayanan, A.~Donnelly, and A.~Rowstron,
``Write Off-Loading: Practical Power Management for Enterprise Storage,"
in \textit{Proceedings of the USENIX Conference on File and Storage Technologies} (FAST), 2008.

\bibitem{Ubuntu}
http://www.ubuntu.com/download

\bibitem{PreemptionGC_Lee}
J.~Lee, Y.~Kim, G.~M.~Shipman, S.~Oral, and J.~Kim,
``Preemptible I/O Scheduling of Garbage Collection for Solid State Drives,"
\emph{IEEE Transactions on Computer-Aided Design of Integrated Circuits and Systems}, vol. 32, no. 2, pp. 247--260, 2013.

\bibitem{PowerHoldUp_Bang}
K.~Bang, K.-I.~Im, D.-G.~Kim, S.-H.~Park, and E.-Y.~Chung,
``Power Failure Protection Scheme for Reliable High-Performance Solid State Disks,"
\emph{IEICE Transactions on Information and Systems}, vol. E96-D, no. 5, pp. 1078--1085, 2013.

\bibitem{Trimming_shi}
L.~Shi, K.~Wu, M.~Zhao, C.~J.~Xue, and E.~H.-M. Sha,
``Retention Trimming for Wear Reduction of Flash Memory Storage Systems,"
in \emph{Proceedings of the Design Automation Conference} (DAC), 2014.

\bibitem{JEDEC_Arrenius}
JEDEC standard,
``Method for Developing Acceleration Models for Electronic Component Failure Mechanisms,"
\emph{JESD91A}, Aug. 2003.

\bibitem{Yu_RFR}
Y.~Cai, Y.~Luo, E.~F.~Haratsch, K.~Mai, and O.~Mutlu,
``Data Retention in MLC NAND Flash Memory: Characterization, Optimization, and Recovery,"
in \emph{Proceedings of the IEEE Symposium on High-Performance Computing Architecture} (HPCA), 2015.

\bibitem{Takeuchi_DRRP}
S.~Tanakamaru, Y.~Yanagihara, and K.~Takeuchi,
``Error-Prediction LDPC and Error-Recovery Schemes for Highly Reliable Solid-State Drives (SSDs),"
\emph{IEEE Journal of Solid-State Circuits}, vol. 48, no. 11, pp. 2920--2933, 2013.

\bibitem{FD_Jeong}
J.~Jeong, Y.~Song, and J.~Kim,
``FlashDefibrillator: A Data Recovery Technique for Retention Failures in NAND Flash Memory,"
in \emph{Proceedings of the IEEE Non-Volatile Memory System and Applications Symposium} (NVMSA), 2015.

\bibitem{Fukuda_RTN}
K.~Fukuda, Y.~Shimizu, K.~Amemiya, M.~Kamoshida, and C.~Hu,
``Random Telegraph Noise in Flash Memories - Model and Technology Scaling,"
in \emph{Proceedings of the IEEE International Electron Devices Meeting} (IEDM), 2007.

\bibitem{Tang_ReadDisturbance}
B.~Tang, C.~Robinson, W.~D.~Zhang, J.~F.~Zhang, R.~Degraeve, P.~Blomme, M.~Toledano-Luque, G.~V.~den bosch, B.~Govoreanu, and J.~V.~Houdt,
``Read and Pass Disturbance in the Programmed States of Floating Gate Flash Memory Cells With High-$k$ Interpoly Gate Dielectric Stacks,"
\emph{IEEE Transactions on Electron Devices}, vol. 60, no. 7, pp. 2261--2267, 2013.

%\bibitem{Refresh_Cai}
%Y.~Cai, G.~Yalcin, O.~Mutlu, E.~F.~Haratsch, A.~Cristal, O.~S.~Unsal, and K.~Mai,
%``Flash Correct-and-Refresh: Retention-Aware Error Management for Increased Flash Memory Lifetime,"
%in \emph{Proceedings of the IEEE International Conference on Computer Design} (ICCD), 2012.

\bibitem{ReadDisturb_Ha}
K.~Ha, J.~Jeong, and J.~Kim,
``An Integrated Approach for Managing Read Disturbs in High-Density NAND Flash Memory,"
accepted for publication in \emph{IEEE Transactions on Computer-Aided Design of Integrated Circuits and Systems}.

\bibitem{SSDAlloc_Badam}
A.~Badam and V. S. Pai,
``SSDAlloc: Hybrid SSD/RAM Memory Management Made Easy,"
in \emph{Proceedings of the USENIX Conference on Networked Systems} (NSDI), 2011.

\bibitem{FlashMainMemory_Jacob}
B.~Jacob, I.~Bhati, M.-T.~Chang, P.~Rosenfeld, J.~Stevens, P.~Tschirhart, Z.~Chishti, S.-L.~Lu, J.~Ang, D.~Resnick, and A.~Rodrigues,
``A Journaled, NAND-Flash Main-Memory System,"
\emph{Technical Report, University of Maryland}, UMD-SCA-2010-12-01, December 2010, updated 2014.

\bibitem{Flash_Facebook}
J.~Taylor,
``Flash at Facebook,"
in \emph{Proceedings of the Flash Memory Summit}, 2013.

\end{thebibliography}
